---
title: "Lecture5_assignment"
output:
  html_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Standard deviation and standard error of the mean
The SD and SEM are different statistical metrics - this exercise will illustrate how they are calculated and related.
* Obtain 1000 samples from a Gaussian distributionwith mean=0 and SD=20 and plot as a histogram.

```{r}
#from lecture 4 assigment: 

gauss_plot <- rnorm(1000, mean =0, sd =20)
hist(gauss_plot)


```


* Calculate the standard deviation of this sample.

```{r}
sd(gauss_plot)
```

* Now, generate the sampling distribution of the mean. Remember - this is simply the distribution of means from sequential samples. Do this sequentially by getting 100 samples of size 3, 10 and 30. Tip: the easiest way to do this will be with a 'for' loop like so:
```
for (i in 1:100) {
  do_interesting_stuff
}
```

```{r}
#Code adapted from: https://www.stat.ncsu.edu/people/ceyhan/courses/st372/ReviewDemo.html


layout(matrix(c(1,2,3), 1,3))


draws1 = sample(gauss_plot, size = 3 * 1000, replace = TRUE) #use the sample function instead of a for loop
draws1 = matrix(draws1, 3)
drawmeans1 = apply(draws1, 2, mean)
hist(drawmeans1, main ="sample size 3")

draws2 = sample(gauss_plot, size = 10 * 1000, replace = TRUE)
draws2 = matrix(draws2, 10)
drawmeans2 = apply(draws2, 2, mean)
hist(drawmeans2, main ="sample size 10")

draws3 = sample(gauss_plot, size = 30 * 1000, replace = TRUE)
draws3 = matrix(draws3, 30)
drawmeans3 = apply(draws3, 2, mean)
hist(drawmeans3, main ="sample size 30")

 

```


* Plot the SD and the three different sampling distributions of the mean in adjacent or superimposed histograms, and calculate their numeric values.

```{r}
#try to make them overlapping -find code online


#layout <-layout(matrix(c(1,2,3,4,5,6), 2,3, byrow=TRUE))
#layout.show(layout)

par(mar= c(2,2,1,1)) 
draws1 = sample(x, size = 3 * 1000, replace = TRUE)
draws1 = matrix(draws1, 3)
drawmeans1 = apply(draws1, 2, mean)
hist(drawmeans1, main ="sample size 3")

par(mar= c(2,2,1,1)) 
draws2 = sample(x, size = 10 * 1000, replace = TRUE)
draws2 = matrix(draws2, 10)
drawmeans2 = apply(draws2, 2, mean)
hist(drawmeans2, main ="sample size 10")

par(mar= c(2,2,1,1)) 
draws3 = sample(x, size = 30 * 1000, replace = TRUE)
draws3 = matrix(draws3, 30)
drawmeans3 = apply(draws3, 2, mean)
hist(drawmeans3, main ="sample size 30")
#hist(x, col ="lightblue")
#arrows(mean(x), 0, mean(x), 200, col ="red")
```


```{r}
#plot gauss_plot next to all three on one plot
hist(drawmeans1, col=rgb(1,0,0,0.5), ylim=c(0,300), xlim=c(-70,70), main= "Overlapping SD and means") #blue
hist(drawmeans2, col=rgb(0,0,1,0.5), add=T) #red
hist(drawmeans3, col="light blue", add=T)
hist(gauss_plot, col =rgb(0.8,0.8,0.8,0.5), add=T) #gray


#code adapted from:https://www.r-bloggers.com/overlapping-histogram-in-r/
```


# Central limit theorem: originating a Gaussian distribution from different probability distributions
Demonstrate the central limit theorem using the following original distributions:
* Gaussian distribution (rnorm) with mean = 0 and SD = 20
* Uniform distribution (runif) with min/max values of 0 and 1
* A binomial distribution (rbinom) with p=0.1
* Chi-square distribution (rchisq) with 3 degrees of freedom

Generate the following samples and plot as a 3x2 layout:
For each one of these, start by characterizing the probability distribution function by drawing 10,000 random samples. Plot the resulting distributions as a histogram.
The code should be something like this:
```
draw 10,000 samples from distribution
plot vector_of_samples as histogram
```

```{r}
#Gaussian:

par(mfcol=c(3,2))

gauss_plot2 <- rnorm(10000, mean =0, sd =20)
hist(gauss_plot2, main ="Gaussian")

#Uniform:
uniform <- runif(10000, min=0, max=1)
hist(uniform, main ="Uniform")

#Binomial:
binomial <- rbinom(10000,size=1, p=0.1 )
hist(binomial, main ="Binomial")

#Chi square:
chisq <- rchisq(10000,3)
hist(chisq, main = "Chi-square")

```

Then, use a 'for' loop to generate 10,000 random drawings of 25 samples each and calculate the mean for each one of these. Plot the resulting distribution of means.

The code should be something like this:
```
for 10,000 iterations 
  draw 10 samples from distribution
  save mean in vector_of_means
end
plot vector_of_means as histogram
```
```{r}

#do the same thing as above: 
par(mfcol=c(3,2))

 
draws1 = gauss_plot2
draws1 = matrix(draws1, 25)
drawmeans1 = apply(draws1, 2, mean)
hist(drawmeans1, main ="Gaussian mean 25 samples")

draws2 = uniform
draws2 = matrix(draws2, 25)
drawmeans2 = apply(draws2, 2, mean)
hist(drawmeans2, main ="Uniform means") 
 
draws3 = binomial
draws3 = matrix(draws3, 25)
drawmeans3 = apply(draws3, 2, mean)
hist(drawmeans3, main ="Binomial means", xlim= c(0,0.5))

draws4 = chisq
draws4 = matrix(draws4, 25)
drawmeans4 = apply(draws4, 2, mean)
hist(drawmeans4, main ="Chi-square means")


#not sure if this method actually works

```


# Central limit theorem: dependency on sample size (OPTIONAL)
The CLT states that the distribution of sample means approximates a normal distribution as the sample size becomes larger. Test this directly: starting with a uniform distribution, plot (as histograms) the distribution of the means from drawing 20, 100 and 5000 samples.

Plotting the vectors of means is useful, but you may need to formally test for normality before you embark on statistical inference using parametric testing. The Shapiro-Wilks test is a statistical test that allows testing for normality in a vector of values. Use it (shapiro.test()) to test whether your samples are normally distributed (p<0.05 means they are).

You can do this easily iterating over the different sample sizes using a 'for' loop. Here's some pseudo-code to get you started:
```
define sample_sizes
for sample in sample_size
  draw n_of_samples from distribution
  plot n_of_samples as histogram
  shapiro_test n_of_samples
end
```