---
title: "Lecture12_assignment"
output:
  html_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Linear regression and associated functions
* Load R data set mtcars using load(mtcars)

```{r}
require(graphics)
require(tidyverse)
#load funtion doesnt work, use data()

data("mtcars")
```

* Next, examine the data using head(), str() or whatever function you prefer.
```{r}
head(mtcars)
```

* Then plot the relationship between hp and mpg using a scatterplot.
```{r}
plot(mtcars$hp, mtcars$mpg)

```


* Run linear regression hp and mpg. Assign the returned lm object to a variable in your workspace. Remember the syntax is lm(y ~ x)
```{r}
linreg<- lm(mtcars$hp ~ mtcars$mpg)

```

* Examine the results of the linear regression using summary(lm_object)
```{r}
summary(linreg)
```

* Verify the fit using plot(lm_object)
```{r}
plot(linreg)
```

* Add the linear regression fit to your original scatterplot using abline(lm_object)
```{r}
#from https://www.statmethods.net/graphs/scatterplot.html

attach(mtcars)
plot(hp, mpg, main="Scatterplot Example",
   xlab="hp ", ylab="Miles Per Gallon ", pch=19)
abline(lm(mpg~hp), col="red") # regression line (y~x)
```

